---
title: "{{ title.test }} for {{ paste0(paste0('`', dvs, '`'), collapse = ',') }}"
author: {{ author }} <{{ email }}>
comment: This file is automatically generate by Shiny-Statistic app (https://statistic.geiser.tech/)
         Author - Geiser C. Challco <geiser@alumni.usp.br>
         
         Shiny-Statistic is distributed in the hope that it will be useful,
         but WITHOUT ANY WARRANTY; without even the implied warranty of
         MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
         GNU General Public License for more details.
         
         You should have received a copy of the GNU General Public License.
         If not, see <https://www.gnu.org/licenses/>.
output:
  github_document:
    toc: true
  word_document:
    toc: true
  html_document:
    toc: true
  pdf_document:
    toc: true
    keep_tex: true
fontsize: 10pt
---

```{r setup, include=FALSE}
## Install and Load Packages
if (!'remotes' %in% rownames(installed.packages())) install.packages('remotes')
if (!"rshinystatistics" %in% rownames(installed.packages())) {
  remotes::install_github("geiser/rshinystatistics")
} else if (packageVersion("rshinystatistics") < "{{ rshinystatistics.version }}") {
  remotes::install_github("geiser/rshinystatistics")
}

wants <- c('ggplot2','ggpubr','rshinystatistics','utils')
has <- wants %in% rownames(installed.packages())
if (any(!has)) install.packages(wants[!has])

library(utils)
library(ggpubr)
library(ggplot2)
library(rshinystatistics)
```

```{r, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
defaultW <- getOption("warn")
options(warn = -1)
```

## Initial Variables and Data

* R-script file: [../code/{{ test }}.R](../code/{{ test }}.R)
* Initial table file: [../data/initial-table.csv](../data/initial-table.csv)
{{ paste0(paste0('* Data for ', dvs, ' [../data/table-for-',dvs,'.csv](../data/table-for-',dvs,'.csv)'), collapse = '\n') }}
* Table without outliers:  [../data/table-without-outliers.csv](../data/table-without-outliers.csv)
* Other data files:  [../data/](../data/)
* Files related to the presented results:   [../results/](../results/)


```{r, include=FALSE}
wid <- "{{ wid }}"
between <- c({{ paste0(paste0('"', between, '"'), collapse = ',') }})
dvs <- c({{ paste0(paste0('"',dvs,'"="',dvs,'"'), collapse = ',') }})

dat <- lapply(dvs, FUN = function(dv) {
  data <- read.csv(paste0("../data/table-for-",dv,".csv"))
  rownames(data) <- data[["{{ wid }}"]]
  return(data)
})
rdat <- dat
```

### Descriptive statistics of initial data

```{r, include=FALSE}
(df <- get.descriptives(dat, dvs, between, include.global = T, symmetry.test = T))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```


```{r, echo=FALSE}
{{
sformula <- paste0(paste0('`',between,'`'), collapse = '*')
box.plot.code  <- paste0(lapply(dvs, FUN = function(dv) {
  paste0("car::Boxplot(`",dv,"` ~ ",sformula,", data = dat[[\"",dv,"\"]], id = list(n = Inf))")
}), collapse = '\n')
box.plot.code
}}
```

## Performing Treatment of Symmetry and Outliers

{{ code.skewness }}

{{ code.outliers }}

```{r}
sdat <- rdat
```

## Saving the Data Without Outliers and Accomplish Symmetry (data for the nonparametric test)

```{r}
ndat <- sdat[[1]]
for (dv in names(sdat)[-1]) ndat <- merge(ndat, sdat[[dv]])
write.csv(ndat, paste0("../data/table-without-outliers.csv"))
```

Descriptive statistics of data used in the nonparametric test

```{r, include=FALSE}
(df <- get.descriptives(sdat, dvs, between))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r, echo=FALSE}
{{
sformula <- paste0(paste0('`',between,'`'), collapse = '*')
box.plot.code  <- paste0(lapply(dvs, FUN = function(dv) {
  paste0("car::Boxplot(`",dv,"` ~ ",sformula,", data = sdat[[\"",dv,"\"]], id = list(n = Inf))")
}), collapse = '\n')
box.plot.code
}}
```

## Computation of Hypothesis Test and Pairwise Comparison

### {{ title.test }}

{{ code.hypothesis }}

{{ code.hypothesis.tbl }}

{{ code.pwc }}

{{ code.pwc.tbl }}

{{ code.plots  }}

### Textual Report

{{ hypothesis.text }}

{{ hypothesis.pwc.text }}

## Tips and References

- Use the site [https://www.tablesgenerator.com](https://www.tablesgenerator.com) to convert the HTML tables into Latex format
