---
title: "{{ title.test }} for {{ as_formula(paste0(dvs,collapse = ','), between, covar = covar, as.character = T) }}"
author: {{ author }} <{{ email }}>
comment: This file is automatically generate by Shiny-Statistic app (https://statistic.geiser.tech/)
         Author - Geiser C. Challco <geiser@alumni.usp.br>
         
         Shiny-Statistic is distributed in the hope that it will be useful,
         but WITHOUT ANY WARRANTY; without even the implied warranty of
         MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
         GNU General Public License for more details.
         
         You should have received a copy of the GNU General Public License.
         If not, see <https://www.gnu.org/licenses/>.
output:
  github_document:
    toc: true
  word_document:
    toc: true
  html_document:
    toc: true
  pdf_document:
    toc: true
    keep_tex: true
fontsize: 10pt
---

```{r setup, include=FALSE}
## Install and Load Packages
if (!'remotes' %in% rownames(installed.packages())) install.packages('remotes')
if (!"rshinystatistics" %in% rownames(installed.packages())) {
  remotes::install_github("geiser/rshinystatistics")
} else if (packageVersion("rshinystatistics") < "{{ rshinystatistics.version }}") {
  remotes::install_github("geiser/rshinystatistics")
}

wants <- c('ggplot2','ggpubr','rshinystatistics','utils')
has <- wants %in% rownames(installed.packages())
if (any(!has)) install.packages(wants[!has])

library(utils)
library(ggpubr)
library(ggplot2)
library(rshinystatistics)
```

```{r, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
defaultW <- getOption("warn")
options(warn = -1)
```

## Initial Variables and Data

* R-script file: [../code/{{ test }}.R](../code/{{ test }}.R)
* Initial table file: [../data/initial-table.csv](../data/initial-table.csv)
{{ paste0(paste0('* Data for ', dvs, ' [../data/table-for-',dvs,'.csv](../data/table-for-',dvs,'.csv)'), collapse = '\n') }}
* Table without outliers and normal distribution of  data: [../data/table-with-normal-distribution.csv](../data/table-with-normal-distribution.csv)
* Other data files: [../data/](../data/)
* Files related to the presented results: [../results/](../results/)


```{r, include=FALSE}
wid <- "{{ wid }}"
{{ ifelse(length(covar) > 0, paste0('covar <- "',covar,'"'), '') }}
between <- c({{ paste0(paste0('"', between, '"'), collapse = ',') }})
dvs <- c({{ paste0(paste0('"', dvs, '"'), collapse = ',') }})
names(dvs) <- dvs

dat <- lapply(dvs, FUN = function(dv) {
  data <- read.csv(paste0("../data/table-for-",dv,".csv"))
  rownames(data) <- data[["{{ wid }}"]]
  return(data)
})
rdat <- dat
sdat <- dat
```

### Descriptive statistics of initial data

```{r, include=FALSE}
{{
if (length(covar) > 0) {
  paste0(c('df <- dat; df[[covar]] <- dat[[1]]',
           '(df <- get.descriptives(df, dvs, between, include.global = T, symmetry.test = T))'), collapse = '\n')
} else {
  paste0(c('(df <- get.descriptives(dat, dvs, between, include.global = T, symmetry.test = T))'), collapse = '\n')
}
}}
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r, echo=FALSE}
{{
sformula <- paste0(paste0('`',between,'`'), collapse = '*')
box.plot.code  <- paste0(lapply(dvs, FUN = function(dv) {
  paste0("car::Boxplot(`",dv,"` ~ ",sformula,", data = dat[[\"",dv,"\"]], id = list(n = Inf))")
}), collapse = '\n')
box.plot.code
}}
```

## Checking of Assumptions

### Assumption: Symmetry and treatment of outliers

#### Applying transformation for skewness data when normality is not achieved

{{ code.skewness }}

#### Dealing with outliers (performing treatment of outliers)

{{ code.outliers }}

### Assumption: Normality distribution of data

#### Removing data that affect normality (extreme values)

```{r}
non.normal <- {{ code.non.normal }}
sdat <- removeFromDataTable(rdat, non.normal, wid)
```

#### Result of normality test in the residual model

```{r, include=FALSE}
{{
if (length(covar) > 0) {
  paste0('(df <- normality.test.by.residual(sdat, dvs, between, c(), covar))')
} else {
  paste0('(df <- normality.test.by.residual(sdat, dvs, between))')
}
}}
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

#### Result of normality test in each group

This is an optional validation and only valid for groups with number greater than 30 observations

```{r, include=FALSE}
(df <- get.descriptives(sdat, dvs, between, include.global = F, normality.test = T))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

**Observation**:

As sample sizes increase, parametric tests remain valid even with the violation of normality [[1](#references)].
According to the central limit theorem, the sampling distribution tends to be normal if the sample is large, more than (`n > 30`) observations.
Therefore, we performed parametric tests with large samples as described as follows: 

- In cases with the sample size greater than 100 (`n > 100`), we adopted a significance level of `p < 0.01`

- For samples with `n > 50` observation, we adopted D'Agostino-Pearson test
that offers better accuracy for larger samples [[2](#references)].

- For samples' size between `n > 100` and `n <= 200`, we ignored the normality test,
and our decision of validating normality was based only in the interpretation of QQ-plots
and histograms because the Shapiro-Wilk and D'Agostino-Pearson tests tend to be too sensitive
with values greater than 200 observation [[3](#references)].

- For samples with `n > 200` observation, we ignore the normality assumption based on the central theorem limit.


{{ linearity.code }}


### Assumption: Homogeneity of data distribution

```{r, include=FALSE}
{{
if (length(covar) > 0) {
  paste0('(df <- homogeneity.test(sdat, dvs, between, c(), covar))')
} else {
  paste0('(df <- homogeneity.test(sdat, dvs, between))')
}
}}
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

## Saving the Data with Normal Distribution Used for Performing {{ title.test }} 

```{r}
ndat <- sdat[[1]]
for (dv in names(sdat)[-1]) ndat <- merge(ndat, sdat[[dv]])
write.csv(ndat, paste0("../data/table-with-normal-distribution.csv"))
```

Descriptive statistics of data with normal distribution

```{r, include=FALSE}
(df <- get.descriptives(sdat, dvs, between))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r, echo=FALSE}
for (dv in dvs) {
  car::Boxplot(`dv` ~ {{ paste0(paste0('`',between,'`'), collapse = '*') }}, data = sdat[[dv]] %>% cbind(dv=sdat[[dv]][[dv]]), id = list(n = Inf))  
}
```

## Computation of {{ title.test }} and Pairwise Comparison

### {{ title.test }}

{{ code.hypothesis }}

{{ code.hypothesis.tbl }}

{{ code.pwc }}

{{ code.pwc.tbl }}

{{ code.emms }}

{{ code.emms.tbl }}

{{ code.plots  }}

### Textual Report

{{ hypothesis.text }}

{{ hypothesis.pwc.text }}

## Tips and References

- Use the site [https://www.tablesgenerator.com](https://www.tablesgenerator.com) to convert the HTML tables into Latex format

- [2]: Miot, H. A. (2017). Assessing normality of data in clinical and experimental trials. J Vasc Bras, 16(2), 88-91.

- [3]: Bárány, Imre; Vu, Van (2007). "Central limit theorems for Gaussian polytopes". Annals of Probability. Institute of Mathematical Statistics. 35 (4): 1593–1621.

```{r include=FALSE, echo=FALSE, purl=FALSE}
options(warn = defaultW)
```
