---
title: "Summary of ANCOVA for {{ paste0(paste0('`', dvs, '`'), collapse = ',') }}"
author: {{ author }} <{{ email }}>
comment: This file is automatically generate by Shiny-Statistic app (https://statistic.geiser.tech/)
         Author - Geiser C. Challco <geiser@usp.br>
         
         Shiny-Statistic is distributed in the hope that it will be useful,
         but WITHOUT ANY WARRANTY; without even the implied warranty of
         MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
         GNU General Public License for more details.
         
         You should have received a copy of the GNU General Public License.
         If not, see <https://www.gnu.org/licenses/>.
output:
  github_document: default
  word_document: default
  html_document: default
  pdf_document:
    keep_tex: true
fontsize: 10pt
---

```{r setup, include=FALSE}
if (!'remotes' %in% rownames(installed.packages())) install.packages('remotes')
if (!"rshinystatistics" %in% rownames(installed.packages())) {
  remotes::install_github("geiser/rshinystatistics")
} else if (packageVersion("rshinystatistics") < "{{ rshinystatistics.version }}") {
  remotes::install_github("geiser/rshinystatistics")
}

wants <- c('ggplot2','ggpubr','rshinystatistics','utils', 'knitr')
has <- wants %in% rownames(installed.packages())
if (any(!has)) install.packages(wants[!has])

library(knitr)
opts_chunk$set(echo = TRUE)

library(utils)
library(ggpubr)
library(ggplot2)
library(rshinystatistics)

defaultW <- getOption("warn")
options(warn = -1)

data <- read.csv("data.csv")
rownames(data) <- data[["{{ wid }}"]]
```

## Initial Data and Script

* Initial data: [data.csv](data.csv)
* R-script file: [ancova.R](ancova.R)

### Setting identificator, dependent and independent variables

```{r}
wid <- "{{ wid }}"
covar <- "{{ covar }}"
between <- c({{ paste0(paste0('"', between, '"'), collapse = ',') }})
dvs <- c({{ paste0(paste0('"', dvs, '"'), collapse = ',') }})
dat <- set_datatable(data, dvs, "var")
```

## Check Assumptions

### Removing outliers from the data

```{r}
outliers <- {{ code.outliers }}
rdat <- remove_from_datatable(dat, outliers, wid, "var")
```

### Normality assumption

**Observation**:

As sample sizes increase, parametric tests remain valid even with the violation of normality <sup>[[1](#references)]</sup>.
According to the central limit theorem, the sampling distribution tends to be normal if the sample is large, more than (`n > 30`) observations. Therefore, we performed parametric tests with large samples as described as follows: 

- In cases with the sample size greater than 30 (`n > 30`), we adopted a significance level of `p < 0.01` instead a significance level of `p < 0.05`.

- In cases with the sample size greater than 100 (`n > 100`), we adopted a significance level of `p < 0.001`

- For samples with `n > 50` observation, we adopted D'Agostino-Pearson test that offers better accuracy for larger samples <sup>[[2](#references)]</sup>.

- For samples' size between `n > 100` and `n <= 200`, we ignored both tests (Shapiro and D'Agostino-Persons), and our decision of normality were based only in the interpretation of QQ-plots and histograms because these tests tend to be too sensitive with values greater than 200 <sup>[[3](#references)]</sup>.

- For samples with `n > 200` observation, we ignore the normality assumption based on the central theorem limit.

#### Applying transformation for skewness data when normality is not achieved

{{ code.skewness }}

#### Removing data that affect normality (extreme values)

```{r}
non.normal <- {{ code.non.normal }}
sdat <- remove_from_datatable(rdat, non.normal, wid, "var")
```

#### Checking normality assumption in the residual model

```r
normality_test_by_res(sdat, dvs, between, c(), covar, dv.var='var')
```

```{r echo=FALSE}
kdf <- normality_test_by_res(sdat, dvs, between, c(), covar, dv.var='var')
kdf$p <- round(kdf$p, 3)
kdf$p[which(kdf$p < 0.001)] <- '<0.001'
kable(kdf, digits = 3)
```

### Linearity assumption

{{ linearity.code }}

### Homogeneity assumption

```r
homogeneity_test(sdat, dvs, between, c(), covar, dv.var='var')
```

```{r echo=FALSE}
kdf <- homogeneity_test(sdat, dvs, between, c(), covar, dv.var='var')
kdf$p <- round(kdf$p, 3)
kdf$p[which(kdf$p < 0.001)] <- '<0.001'
kable(kdf, digits = 3)
```

Homogeneity may be ignored employing Wilcox's method (not jet available in rshiny-statistics).

## Computation ANCOVA and Pairwise comparison

### ANCOVA test

```r
aov <- ancova.test(sdat, dvs, between, covar, "{{ type }}", "{{ effect.size }}", "var")
(get.ancova.table(aov))
```

```{r echo=FALSE}
aov <- ancova.test(sdat, dvs, between, covar, "{{ type }}", "{{ effect.size }}", "var")
cnames <- c("var", "Effect", "DFn", "DFd", "SSn", "SSd", "F", "p", "{{ effect.size }}", "p.signif")
kdf <- get.ancova.table(aov)[,cnames]
kdf$p <- round(kdf$p, 3)
kdf$p[which(kdf$p < 0.001)] <- '<0.001'
kable(kdf, digits = 3)
```

### Pairwise comparison showing only significant differences

```r
pwc <- ancova.pwc(sdat, dvs, between, covar, "{{ p.adjust.method }}", "var")
(get.ancova.pwc.table(pwc, only.sig = T))
```

```{r echo=FALSE}
pwc <- ancova.pwc(sdat, dvs, between, covar, "{{ p.adjust.method }}", "var")
cnames <- c("var", between, "group1", "group2","estimate", "se","df","statistic","p", "p.adj","p.adj.signif")
kdf <- get.ancova.pwc.table(pwc, only.sig = T)[,cnames]
kdf$p <- round(kdf$p, 3)
kdf$p[which(kdf$p < 0.001)] <- '<0.001'
kdf$p.adj <- round(kdf$p.adj, 3)
kdf$p.adj[which(kdf$p.adj < 0.001)] <- '<0.001'
kable(kdf, digits = 3)
```

## Estimated Marginal Means and ANCOVA Plots 

```r
get.ancova.emmeans.with.ds(pwc, sdat, dvs, between, "common", "var")
```

```{r echo=FALSE}
kdf <- get.ancova.emmeans.with.ds(pwc, sdat, dvs, between, "common", "var")
cnames <- c("var",between,covar,"n","emmean","se.emms","conf.low","conf.high","mean","median","sd","ci")
kdf <- kdf[,cnames]
kable(kdf, digits = 3)
```

{{ ancova.plots  }}

## Textual Report

{{ ancova.text }}

{{ ancova.pwc.text }}

## References

<sup>[1]</sup>: Ghasemi, A., & Zahediasl, S. (2012). Normality tests for statistical analysis: a guide for non-statisticians. International journal of endocrinology and metabolism, 10(2), 486.

<sup>[1]</sup>: Miot, H. A. (2017). Assessing normality of data in clinical and experimental trials. J Vasc Bras, 16(2), 88-91.


```{r include=FALSE}
options(warn = defaultW)
```
