---
title: "ANOVA for {{ paste0('`',dv,'` ~ ',paste0(paste0('`',between,'`'), collapse='*')) }}"
author: {{ author }} <{{ email }}>
comment: This file is automatically generate by Shiny-Statistic app (https://statistic.geiser.tech/)
         Author - Geiser C. Challco <geiser@usp.br>
         
         Shiny-Statistic is distributed in the hope that it will be useful,
         but WITHOUT ANY WARRANTY; without even the implied warranty of
         MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
         GNU General Public License for more details.
         
         You should have received a copy of the GNU General Public License.
         If not, see <https://www.gnu.org/licenses/>.
output:
  github_document: default
  word_document: default
  html_document: default
  pdf_document:
    keep_tex: true
fontsize: 10pt
---

```{r setup, include=FALSE}
if (!'remotes' %in% rownames(installed.packages())) install.packages('remotes')
if (!"rshinystatistics" %in% rownames(installed.packages())) {
  remotes::install_github("geiser/rshinystatistics")
} else if (packageVersion("rshinystatistics") < "{{ rshinystatistics.version }}") {
  remotes::install_github("geiser/rshinystatistics")
}

wants <- c('ggplot2','ggpubr','rshinystatistics','car','stats','rstatix','utils','dplyr','knitr')
has <- wants %in% rownames(installed.packages())
if (any(!has)) install.packages(wants[!has])

library(knitr)
opts_chunk$set(echo = TRUE)

library(utils)
library(ggpubr)
library(ggplot2)
library(rstatix)
library(rshinystatistics)

defaultW <- getOption("warn")
options(warn = -1)

dat <- read.csv("data.csv")
rownames(dat) <- dat[["{{ wid }}"]]
```

## Initial Data and R-Script

* Initial data: [data.csv](data.csv)
* R-script file: [factorialAnova.R](factorialAnova.R)

## Check Assumptions

### Identifying outliers

Outliers tend to increase type-I error probability, and they decrease the calculated statistic in parametric tests resulting in a lower chance of reject the null hypothesis.

* Identified outliers using rstatix

```{r}
identify_outliers(dplyr::group_by(dat, {{ paste0(paste0('`',between,'`'), collapse = ',') }}), `{{ dv }}`)
```

* Identified outliers through Boxplots

```{r}
car::Boxplot(`{{ dv }}` ~ {{ paste0(paste0('`',between,'`'), collapse = '*') }}, data = dat, id = list(n = Inf))
```

### Removing outliers from the data

```{r}
outliers <- c({{ paste0(paste0('"',outlier.ids,'"'),collapse = ',') }})
rdat <- dat[!dat[["{{ wid }}"]] %in% outliers,]
```

### Normality assumption

**Observation**:

As sample sizes increase, parametric tests remain valid even with the violation of normality <sup>[[1](#references)]</sup>.
According to the central limit theorem, the sampling distribution tends to be normal if the sample is large, more than (`n > 30`) observations. Therefore, we performed parametric tests with large samples as described as follows: 

- In cases with the sample size greater than 30 (`n > 30`), we adopted a significance level of `p < 0.01` instead a significance level of `p < 0.05`.

- In cases with the sample size greater than 100 (`n > 100`), we adopted a significance level of `p < 0.001`

- For samples with `n > 50` observation, we adopted D'Agostino-Pearson test that offers better accuracy for larger samples <sup>[[2](#references)]</sup>.

- For samples' size between `n > 100` and `n <= 200`, we ignored both tests (Shapiro and D'Agostino-Persons), and our decision of normality were based only in the interpretation of QQ-plots and histograms because these tests tend to be too sensitive with values greater than 200 <sup>[[3](#references)]</sup>.

- For samples with `n > 200` observation, we ignore the normality assumption based on the central theorem limit.

#### Applying transformation for skewness data when normality is not achieved

{{ code.skewness }}

#### Removing data that affect normality (extreme values)

```{r}
non.normal <- c({{ paste0(paste0('"',non.normal.ids,'"'),collapse = ',') }})
sdat <- rdat[!rdat[["{{ wid }}"]] %in% non.normal,]
rownames(sdat) <- sdat[["{{ wid }}"]]
```

#### Checking normality assumption in the residual model

```r
normality_test_by_res(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}))
```

```{r echo=FALSE}
kdf <- normality_test_by_res(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}))
kdf$p <- round(kdf$p, 3)
kdf$p[which(kdf$p < 0.001)] <- '<0.001'
kable(kdf, digits = 3)
```

* The QQ plot used to evaluate normality assumption

```{r}
res <- stats::residuals(stats::lm(`{{ dv }}` ~ {{ paste0(paste0('`',between,'`'), collapse='*') }}, data = sdat))
car::qqPlot(res, ylab = "sample")
```


#### Checking normality assumption in the groups

```r
normality_test_per_group(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}))
```

```{r echo=FALSE}
kdf <- normality_test_per_group(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}))
kdf$p <- round(kdf$p, 3)
kdf$p[which(kdf$p < 0.001)] <- '<0.001'
kable(kdf, digits = 3)
```


* The QQ plot used to evaluate normality assumption in groups

{{ group.qqplot }}


### Homogeneity assumption

```r
homogeneity_test(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}))
```

```{r echo=FALSE}
kdf <- homogeneity_test(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}))
kdf$p <- round(kdf$p, 3)
kdf$p[which(kdf$p < 0.001)] <- '<0.001'
kable(kdf, digits = 3)
```

Homogeneity may be ignored in one-way ANOVA or employing Wilcox's method (not jet available in rshiny-statistics).

## Computation ANOVA and Pairwise comparison

### ANOVA test

```r
aov <- get.anova.test(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}), type={{ type }}, effect.size="{{ effect.size }}")
(get.anova.table(aov))
```

```{r echo=FALSE}
aov <- get.anova.test(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}), type={{ type }}, effect.size="{{ effect.size }}")
cnames <- c("Effect", "DFn", "DFd", "SSn", "SSd", "F", "p", "{{ effect.size }}", "p.signif")
kdf <- get.anova.table(aov)[,cnames]
kdf$p <- round(kdf$p, 3)
kdf$p[which(kdf$p < 0.001)] <- '<0.001'
kable(kdf, digits = 3)
```

### Pairwise comparison showing all comparisons

```r
pwc <- get.anova.pwc(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}), p.adjust.method="{{ p.adjust.method }}")
(get.anova.pwc.table(pwc))
```

```{r echo=FALSE}
pwc <- get.anova.pwc(sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse = ',') }}), p.adjust.method="{{ p.adjust.method }}")
cnames <- c(c({{ paste0(paste0('"',between,'"'), collapse = ',') }}), "group1", "group2","estimate", "se","df","statistic","p", "p.adj","p.adj.signif")
kdf <- get.anova.pwc.table(pwc)[,cnames]
kdf$p <- round(kdf$p, 3)
kdf$p[which(kdf$p < 0.001)] <- '<0.001'
kdf$p.adj <- round(kdf$p.adj, 3)
kdf$p.adj[which(kdf$p.adj < 0.001)] <- '<0.001'
kable(kdf, digits = 3)
```

## Estimated Marginal Means and ANOVA Plots 

```r
get.anova.emmeans.with.ds(pwc, sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse=',') }}), "common")
```

```{r echo=FALSE}
kdf <- get.anova.emmeans.with.ds(pwc, sdat, "{{ dv }}", c({{ paste0(paste0('"',between,'"'), collapse=',') }}), "common")
cnames <- c("var",c({{ paste0(paste0('"',between,'"'), collapse = ',') }}),"n","emmean","se.emms","conf.low","conf.high","mean","median","sd","ci")
kdf <- kdf[,cnames]
kable(kdf, digits = 3)
```

{{ anova.plots  }}


## References

<sup>[1]</sup>: Ghasemi, A., & Zahediasl, S. (2012). Normality tests for statistical analysis: a guide for non-statisticians. International journal of endocrinology and metabolism, 10(2), 486.

<sup>[1]</sup>: Miot, H. A. (2017). Assessing normality of data in clinical and experimental trials. J Vasc Bras, 16(2), 88-91.


```{r include=FALSE}
options(warn = defaultW)
```
