---
title: "Wilcoxon test for {{ paste0(paste0('`', dvs, '`'), collapse = ',') }} with `{{ iv }}`"
author: {{ author }} <{{ email }}>
comment: This file is automatically generate by Shiny-Statistic app (https://statistic.geiser.tech/)
         Author - Geiser C. Challco <geiser@alumni.usp.br>
         
         Shiny-Statistic is distributed in the hope that it will be useful,
         but WITHOUT ANY WARRANTY; without even the implied warranty of
         MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
         GNU General Public License for more details.
         
         You should have received a copy of the GNU General Public License.
         If not, see <https://www.gnu.org/licenses/>.
output:
  github_document:
    toc: true
  word_document:
    toc: true
  html_document:
    toc: true
  pdf_document:
    toc: true
    keep_tex: true
fontsize: 10pt
---

```{r setup, include=FALSE}
## Install and Load Packages
if (!'remotes' %in% rownames(installed.packages())) install.packages('remotes')
if (!"rshinystatistics" %in% rownames(installed.packages())) {
  remotes::install_github("geiser/rshinystatistics")
} else if (packageVersion("rshinystatistics") < "{{ rshinystatistics.version }}") {
  remotes::install_github("geiser/rshinystatistics")
}

wants <- c('ggplot2','ggpubr','rshinystatistics','utils')
has <- wants %in% rownames(installed.packages())
if (any(!has)) install.packages(wants[!has])

library(utils)
library(ggpubr)
library(ggplot2)
library(rshinystatistics)
```

```{r, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
defaultW <- getOption("warn")
options(warn = -1)
```

## Initial Variables and Data

* R-script file: [wilcoxon.R](wilcoxon.R)
{{ paste0(paste0('* Data for ', dvs, ' [data-',dvs,'.csv](data-',dvs,'.csv)'), collapse = '\n') }}
* Table without outliers:  [../data/table-without-outliers.csv](../data/table-without-outliers.csv)
* Other data files:  [../data/](../data/)
* Files related to the presented results:   [../results/](../results/)


```{r, include=FALSE}
wid <- "{{ wid }}"
iv <- c({{ paste0(paste0('"', iv, '"'), collapse = ',') }})
dvs <- c({{ paste0(paste0('"', dvs, '"'), collapse = ',') }})
names(dvs) <- dvs

dat <- lapply(dvs, FUN = function(dv) {
  data <- read.csv(paste0("../data/table-for-",dv,".csv"))
  rownames(data) <- data[["{{ wid }}"]]
  return(data)
})
rdat <- dat
sdat <- dat
```

### Descriptive statistics of initial data

```{r, include=FALSE}
(df <- descriptive_statistics(dat, dvs, iv, include.global = T, symmetry.test = T))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```


```{r, echo=FALSE}
for (dv in dvs) {
  car::Boxplot(`dv` ~ {{ paste0(paste0('`',iv,'`'), collapse = '*') }}, data = dat[[dv]] %>% cbind(dv=dat[[dv]][[dv]]), id = list(n = Inf))  
}
```

## Checking of Assumptions

### Assumption: Symmetry and treatment of outliers

#### Applying transformation for skewness data when normality is not achieved in any dependent variables or covariance

{{ code.skewness }}

#### Dealing with outliers (performing treatment of outliers)

{{ code.outliers }}

### Assumption: Normality distribution of data

#### Removing data that affect normality (extreme values)

```{r}
non.normal <- {{ code.non.normal }}
sdat <- remove_from_datatable(rdat, non.normal, wid)
```

## Removing outliers from the data

```{r}
outliers <- {{ code.outliers }}
sdat <- remove_from_datatable(dat, outliers, wid, "var")
```


## Saving the Data without outliers

```{r}
ndat <- sdat[[1]]
for (dv in names(sdat)[-1]) ndat <- merge(ndat, sdat[[dv]])
write.csv(ndat, paste0("../data/table-without-outliers.csv"))
```

Descriptive statistics of data with normal distribution

```{r, include=FALSE}
(df <- descriptive_statistics(sdat, dvs, iv, include.global = T, normality.test = T))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r, echo=FALSE}
for (dv in dvs) {
  car::Boxplot(`dv` ~ {{ paste0(paste0('`',iv,'`'), collapse = '*') }}, data = sdat[[dv]] %>% cbind(dv=sdat[[dv]][[dv]]), id = list(n = Inf))  
}
```

## Computation of Wilcoxon and Pairwise Comparison

### Wilcoxon

```{r, include=FALSE}
(res <- wilcoxon_test(sdat, dvs, iv, "{{ alternative }}", as.list=T))
```

```{r, echo=FALSE, purl=FALSE}
kable(res$wilcoxon.test[,c(".y.", "group1", "group2", "n1", "n2","statistic", "estimate",
  "conf.low", "conf.high", "effsize", "magnitude", "p","p.signif")], digits = 3)
```


## Descriptive Statistic and Plots 

```{r, include=FALSE}
(df <- descriptive_statistics(sdat, dvs, iv, "common", "var"))
```

```{r echo=FALSE, purl=FALSE}
cnames <- c("variable",iv,"n","median","mean","min","max","iqr","sd")
kable(df[,cnames], digits = 3)
```

{{ wtest.plots  }}

## Textual Report

{{ wilcoxon.text }}


## Tips and References

- Use the site [https://www.tablesgenerator.com](https://www.tablesgenerator.com) to convert the HTML tables into Latex format

